Management and effectiveness of community-based marine reserves in small-scale fisheries

### 2018-06-08 docs/paper/Villasenor-Derbez.Rmd
Made changes to manuscript body, just missing details on discussion

---
title: "Enabling conditions for effective community-based marine reserves in small-scale fisheries"
rtitle: "Marine reserve effectiveness"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: "Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities. A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits). In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing. While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves. In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico. By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness. Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness. For example, the most effective reserve had strong governance structures accompanied with low environmental variability. Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa. Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making."
keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

@@

**Last_update**: `r Sys.Date()`

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to manage the spatial distribution of fishing effort and recover stocks is through the implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}.

Marine reserve science has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change mitigation \citep{roberts_2017-J9}, and protection from environmental variability \citep{micheli_2012-EU}. However, there is considerably less literature focusing on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely provide a holistic view of the social-ecological system \citep{lpezangarita_2014}.

While ecological factors (*e.g.* habitat representation, initial state of protection, connectivity to other protected areas) can determine the success of a MR, their effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed. Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs. For example, age of a reserve (*i.e.* time since its implementation) and size were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}. In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers. At a global level, \citet{edgar_2014-UO} indicate that enforcement, age, size, and isolation were important factors determining effectiveness of the reserves.

Marine Reserves in Mexico have been commonly implemented as "core zones" within Biosphere Reserves that are administered by the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP). While CONANP has made efforts to have a participatory process, the implementation of these zones is still characterized by top--down approaches. This motivated Civil Society Organizations (CSOs) to work with coastal communities to implement community--based marine reserves \citep{uribe_2010-u2}, which are usually established within a TURFs; thus making them TURF--reserves \citep{afflerbach_2014-HP}. This bottom--up approach allows fishers to design their own reserves, which  increases compliance and self--enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However, these reserves still lack legal recognition, making them vulnerable to poaching. In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community--based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*, FR). Since then, **39** FR have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.

We combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg} to provide a comprehensive ecological and socioeconomic evaluation of community-based marine reserves in three coastal communities in Mexico. The objective of this work is twofold: i) Provide the first evaluation of community--based marine reserves in Mexico, and ii) provide a comprehensive evaluation of the social--ecological system to identify how socioeconomic and governance characteristics relate to ecological effectiveness.

@@

We focus our evaluation in three coastal communities from the Pacific coast of Baja California and the Mesoamerican Reef System (Fig \ref{fig:map}). Fishing activities in all communities are assambled around fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs). Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests (*Macrocystis pyrifera*) and rocky reefs are the predominant habitats. The island is home to a the *Sociedad Cooperative de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, spiny lobster (*Panulirus interruptus*) represents an important fishery. However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M. undosa*), and abalone (*Haliotis spp*, until 2010) are also important sources of income. In 2006, the community decided to implement two community--based marine reserves within their fishing grounds, seeking to recover depleted stocks of invertebrate species (mainly lobster and abalone). Until today, these reserves are yet to be legally recognized as Fishing Refugia but count with full support from the community.

The other two communities are Maria Elena (ME) and Punta Herrero (PH; Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems. ME is a fishing camp --visited intermittently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*). PH is also home to a fishing cooperative (*SCPP José María Azcorra*). The main source of income to both communities is the Caribbean spiny lobster (*Panulirus argus*). These communities also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae). ME established eight marine reserves in 2012, and PH established four marine reserves in 2013 and an additional community-based (*i.e.* not legally recognized). All these reserves are legally recognized as Fishing Refugia.

@@

We use three main sources of information to evaluate these reserves. Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO "Comunidad y Biodiversidad" ([COBI](www.cobi.org.mx)). These monitorings record richness and abundances of fish and invertebrate species in the reserves and control sites. Size structures are also collected in fish census. We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had the same probability of being selected as reserves during the design phase. From all the reserves in these three communities, we use the ones that have data for reserve and control sites before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were surveyed annually from at least one year before implementation until 2016. Table \ref{table:com_sum} shows a summary of sampling effort and year of implementation of the analyzed reserves.

@@

```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A). Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula."}
# Define reference map

@@

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014. This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} via Eq \ref{eqn:cpi}.

@@

We use a set of governance indicators to analyze the governance structures of each cooperative \citep{leslie_2015-na}. The indicators (Table \ref{table:gov_ind}) focus on the resource system (four indicators), governance system (seven indicators), resource units (three indicators) and actors (three indicators). These were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2012} and based on the authors' experience and knowledge of the communities.

@@

We evaluate the effect thar marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}). Recall that reserves were implemented to protect lobster and other benthic invertebrates. However, we also contemplate fish species that are also targeted in the off-season and might receive collateral protection from the reserves.

```{r}
tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
           Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
       Units = c("org / $m^2$", "org / $m^2$", "org / $m^2$", "org / $m^2$", "M MXP", "Metric Tones")) %>% 
  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

```{r}
read.csv(here::here("data", "gov_ind.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_ind}Indicators used for the operationalization of the SES framework \\citep{leslie_2015-na}", col.names = c("Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero"), format = "latex") %>% 
  kableExtra::column_spec(1, width = "10em") %>% 
  kableExtra::column_spec(2, width = "8em") %>% 
  kableExtra::column_spec(3, width = "8em") %>% 
  kableExtra::column_spec(4, width = "8em") %>% 
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(6 , bold = T) %>% 
  kableExtra::row_spec(14 , bold = T) %>% 
  kableExtra::row_spec(18 , bold = T)
```

Biological indicators are analyzed with a difference--in--differences analysis (Eq \ref{eqn:reg_bio}), which allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments \citep{moland_2013-VP,Villasenor-Derbez_2018}. The analysis is performed with a multiple linear regression of the form:

@@

Where year--fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The interaction term $\lambda_{it} Year_t\times Zone_i$ represents represent the mean change in the indicator inside the reserve, for year $t$, with respect to the first year of evaluation in the control site (See Table \ref{table:com_sum}). When evaluating biomass and abundances, we include species--fixed effects ($\sigma_j$).

@@

\begin{equation}
I_i = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jComm_j +\epsilon
\label{eqn:soc_reg}
\end{equation}

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jComm$ captures community-level fixed-effects. This approach allows for a causal attribution of the observed changes to the reserve. All model coefficients were estimated via ordinary least--squares and heteroskedastic--robust standard errors \citep{zeileis_2004-7n}.

These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the reserve site at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively).

@@

The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community. Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size). We also provide an overview of the governance settings of each community, and discuss how these dimensions might be related to the effectiveness of the reserves.

@@

All indicators show idiosyncratic responses through time for each community. Figure \ref{fig:indicators}A shows how lobster densities increase inside the reserve for Isla Natividad an Punta Herrero within the first years, but the effect is aroded through time. However, these effects are in the order of 0.2 extra organisms $m^{-2}$ and are not significantly different from zero ($p > 0.05$). The only case when lobster densities showed a significant positive effect ($p < 0.1$) was for Isla Natividad on the sixth year (*i.e.* 2012), a year after hypoxia events described in \citet{micheli_2012-EU} caused mass mortality of organisms. Biomass was only evaluated for fish data (Fig \ref{fig:indicators}B), and no changes were observed ($p > 0.1$), Punta Herrero even showed negative insignificant trends in the order of 0.01. Invertebrate and fish densities (\ref{fig:indicators}C-D) show similar patterns of lack of effectiveness with effect sizes oscillating around zero. Full tables with model coefficients are presented in the supplementary materials (**S1 Table**, **S2 Table**, **S3 Table**).

@@

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}). For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well. However, effect sizes do not change after the implementation of the reserve. Again, the negative coeficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The trhee years of post-implementation data for Maria Elena do not show a significant effect of the reserve. Isla Natividad shows higher revenues after the iumplementation of the reserve, as compared to the control communities. However, these changes are also not ignificantly different and present an increased variation. All regression coefficients for each community and indicator are presented in **S4 Table**.

@@

Although we have little information on the social dimension of these fisheries, using the SES framework indicators (Table \ref{table:gov_ind}), we can analyze the performance of each governance system (Table \ref{table:gov_res}).

@@

```{r}
read.csv(here::here("data","gov_res.csv")) %>% 
  knitr::kable(caption = "\\label{table:gov_res}Analysis of the fishing cooperatives based on the Social-Ecological systems framework \\citep{mcginnis_2014}.",
               col.names = c("", "Indicator", "Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  kableExtra::column_spec(1, width = "10em") %>%
  kableExtra::column_spec(2, width = "10em") %>%
  kableExtra::row_spec(1 , bold = T) %>% 
  kableExtra::row_spec(7 , bold = T) %>% 
  kableExtra::row_spec(15 , bold = T) %>% 
  kableExtra::row_spec(19 , bold = T)
```

@@

theory tells us more than 5 years are needed, so it;s not surprise for PH and ME. But IN didnt make it either.

Our results indicated idiosyncratic biological effects of the reserves across communities and indicators, with a combination of positive and negative effects. However, many of these effects were not statistically significant, indicating no effect of the reserve. The socioeconomic indicators pertaining to landings and revenues showed little or no change after reserve implementation. While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and as a double-check. ll indicators presented a similar pattern. The lack of expected effectiveness poses the question, why do these communities continue to support the reserves? While not in the main scope of this paper, understanding the social-ecological context in which these communities and their reserves might provide insight to answer this question. Here, we discuss plausible explanations to lack of effectiveness observed based on literature and our social-ecological system analysis. We also discuss potential shortcomings in our analysis, and provide management recommendations to improve reserve effectiveness.

Our approach to evaluate the temporal and spatial changes of each indicator provides a more robust measure of reserve effectiveness. Some works have solely focused on an inside-outside comparison of indicators \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do not address temporal variability \citep{depalma_2018}. Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}. By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect

@@

@@

@@

\paragraph*{S1 Figure}
\label{S1_Figure}
Timeseries of indicators for IN

\paragraph*{S2 Figure}
\label{S2_Figure}
Timeseries of indicators for ME

\paragraph*{S3 Figure}
\label{S3_Figure}
Timeseries of indicators for PH

>>>>

---
title: "Management and effectiveness of community-based marine reserves in small-scale fisheries"
rtitle: "Community-based marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Álvin Suarez-Castillo$^{2}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: "Al finalizar revisiones"
keywords: "Marine Protected Areas, Marine Conservation, Small-Scale Fisheries, Citizen Science, Mexico, Social-Ecological Systems"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

<!-- Old Abstract -->
<!-- Coastal marine ecosystems provide livelihoods for small-scale fishers and coastal communities around the world. However, overfishing and unsustainable fishing practices threaten the marine environment and jeopardize the wellbeing of coastal communities. A common approach to protect the environment and recover overexploited stocks is to implement no-take marine reserves (areas where all extractive activities are off-limits). In small-scale fisheries, these are sometimes implemented as community-based reserves, where a group of fishers collectively agree to close an area to fishing. While we know that reductions in fishing effort are followed by a series of ecological benefits (increased biomass, abundance, and species diversity), we do not fully understand how environmental and governance dynamics influence the conservation and fisheries benefits of community-based marine reserves. In this work, we evaluate the ecological outcomes of four reserves established by three coastal communities in temperate and tropical ecosystems of Mexico. By combining causal inference techniques with an operationalization of the social-ecological systems framework, we identify the environmental and social conditions that enable reserve effectiveness. Our results show a strong interaction between environmental variation and community organization, which influences reserve effectiveness. For example, the most effective reserve had strong governance structures accompanied with low environmental variability. Thus, even when a community is well organized (and reserves are well enforced), environmental variation can hinder the benefits of a reserve, and vice versa. Our results are particularly relevant under present changing climate conditions, as they can better inform management and decision making. -->

@@

**Titulos alternativos**

"Community-based marine reserves in small-scale fisheries"

"Effectiveness of community-based marine reserves: lessons for their management and implementation" (no me encanta)

Algun otro?

**Last update**: `r Sys.Date()`

@@

Marine ecosystems around the world sustain significant impacts due to overfishing and unsustainable fishing practices \citep{halpern_2008-dK,worm_2006-IB,pauly_2005-qV}. A common approach to manage the spatial distribution of fishing effort to recover stocks and preserve biodiversity is through the implementation of marine reserves (MRs). These areas allow bounded populations to recover by limiting all extractive activities \citep{halpern_2002}. The science of MRs has largely focused on understanding the ecological effects of these areas, which include increased biomass, richness, and densities of organisms within the protected regions, climate change mitigation, and protection from environmental variability \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69,roberts_2017-J9,micheli_2012-EU}. Modelling studies show that the fishery benefits of marine reserves depend on initial stock status and the type of management under which the fishery operates \citep{hilborn_2006}, as well as reserve size and the amount of larvae exported from these \citep{krueck_2017-J1}. Other research has focused on the relationship between socioeconomic and governance structures and their relationship to ecological effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. However, few studies simultaneously evaluate MRs from all these perspectives (*e.g.* \citet{lpezangarita_2014}).

While ecological factors like habitat representation within the MR or connectivity to other areas can determine the success of a MR, its effectiveness also depends on the socioeconomic and governance settings under which they are implemented and managed. Literature shows that many non-ecological characteristics can play an equally important role in the effectiveness of MRs. For example, age of a reserve (*i.e.* time since its implementation) and size were key to the effectiveness of MRs in Palau \citep{friedlander_2017-oI}. In the Mediterranean, \citet{difranco_2016-Xw} identify that surveillance and enforcement, presence of a management plan, and involvement of fishers in management and decision--making along with promotion of sustainable fishing practices were the key factors that increased stock health and income to fishers. At a global level, enforcement, age, size, and isolation are important factors that determine the effectiveness of the reserves \citep{edgar_2014-UO}. However, MRs represent part of a more complex social-ecological systems governed by the interaction between humans and the environment. In this work we evaluate community-based marine reserves in Mexico using a triple bottom line approach by using ecological, socioeconomic, and governance indicators that provide a holistic view of the reserves \citep{halpern_2013}.

There are three main approaches to implement MRs, each of which have its pros and cons. We describe them in the context of Mexican MRs, but argue that at least the first two apply elsewhere. MRs were historically implemented and managed by a government agency, in this case the National Commission of Protected Areas (*Comisión Nacional de Áreas Marinas Protegidas*, CONANP). While CONANP has made efforts to have a participatory process, the implementation of these zones is still mainly a top-down process. A second approach is the implementation of community-based marine reserves within areas of exclusive access (*i.e.* TURFs), thus making them TURF-reserves \citep{afflerbach_2014-HP}. Community-based spatial closures ocurr in other places, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. This bottom-up approach increases compliance and self-enforcement \citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However, these lack legal recognition and rely on the exclusive access granted by the TURF, making it impossible to enforce without one. In an effort to provide a legal framework for these reserves, Civil Society Organizations (CSOs) served as the bridge between fishers and government. In 2014, a new norm \citep{nom} allowed fishers to request the legal recognition of a community-based reserve under the name of "Fishing Refugia" (*Zona de Refugio Pesquero*). These can be implemented as temporal or partial eservs, which can protect one, some, or all resources within them. Since then, **39** of these have been implemented along the Pacific, Gulf of California, and Mexican Caribbean coastlines, but no formal evaluation of their effectiveness has taken place.

This work combines causal inference techniques and the social-ecological systems framework to provide a holistic evaluation of community-based marine reserves in three coastal communities in Mexico. The objective of this work is twofold. First, provide a triple bottom line evaluation of the effectiveness of community-based marine reserves that can inform similar processes in other countries. And second, perform the first formal evaluation of Fishing Refugia in Mexico and identify areas where improvement or adjustment might result in increased effectiveness. On both cases, we draw from the lessons learned and provide management recommendations to ensure or improve the effectiveness of community-based marine reserves.

@@

We evaluate community-based marine reserves from three coastal communities located in the Pacific coast of Baja California and the Mexican Caribbean (Fig \ref{fig:map}). All communities are organized as fishing cooperatives that hold Territorial Use Rights for Fisheries (TURFs). Isla Natividad (IN) lies west of the Baja California Peninsula (Fig \ref{fig:map}B), where kelp forests and rocky reefs are the predominant habitats. The island is home to the *Sociedad Cooperativa de Producción Pesquera (SCPP) Buzos y Pescadores de la Baja California*, whose main resource by value is the spiny lobster (*Panulirus interruptus*). However, other resources like finfish (yellow-tail jack, *Seriola lalandi*), sea cucumber (*Parastichopus parvimensis*), red sea urchin (*Mesocentrotus franciscanus*), snail (*Megastraea turbanica* y *M. undosa*), and abalone (*Haliotis spp*) are also important sources of income. In 2006, the community decided to implement two community-based marine reserves within their fishing grounds to protect commercially important invertebrate species (mainly lobster and abalone). Until today, these reserves are yet to be legally recognized as Fishing Refugia but count with full support from the community.

The other two communities are Maria Elena (ME) and Punta Herrero (PH; Fig \ref{fig:map}C) in the Yucatan Peninsula, where coral reefs and mangroves are the representative coastal ecosystems. ME is a fishing camp --visited intermitently during the fishing season-- belonging to the Cozumel fishing cooperative (*SCPP Cozumel*); PH is home to the *SCPP José María Azcorra* cooperative. Their main fishery is the Caribbean spiny lobster (*Panulirus argus*), but they also target finfish in the off season, mainly snappers (Lutjanidae) and groupers (Serranidae). ME established eight marine reserves in 2012, and PH established four marine reserves in 2013 and an additional community-based (*i.e.* not legally recognized). All these reserves are legally recognized as Fishing Refugia.

@@

We use three main sources of information to evaluate these reserves. Ecological data come from the annual ecological monitoring of reserve and control areas, carried out by members from each community and personnel from the Mexican CSO *Comunidad y Biodiversidad* ([COBI](www.cobi.org.mx)). Trained divers record richness and abundances of fish and invertebrate species in the reserves and control sites. Size structures are also collected for fish evaluations. We define control sites as regions with habitat characteristics similar to the corresponding reserves, and that presumably had a similar probability of being selected as reserves during the design phase. We focus our evaluation for sites where data are avaialable for reserve and control sites, before and after the implementation of the reserve. This provides us with a Before-After-Control-Impact (*i.e.* BACI) sampling design that allows us to capture and control for temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI designs and causal inference techniques have proven effective to evaluate marine reserves, as they allow us to causally attribute observed changes to the intervention \citep{moland_2013-VP,Villasenor-Derbez_2018}. All sites were surveyed annually, and at least once before implementation of the reserves. Table \ref{table:com_sum} shows a summary of the reserves included in this study.

@@

```{r, fig.cap = "\\label{fig:map}Location of the three coastal communities studied (A). Isla Natividad (B) is located off the Baja California Peninsula, Maria Elena (C) and Punta Herrero (D) are located in the Yucatan Peninsula. Blue polygons represent the TURFs, and red polygons the marine reserves."}
# Define reference map

@@

Socioeconomic data come from landing receipts reported to the National Commission for Aquaculture and Fisheries (*Comisión Nacional de Acuacultura y Pesca*; CONAPESCA). Data contain monthly lobster landings (Kg) and value (MXP) from 2000 to 2014 for cooperatives with and without marine reserves (**Fig S1**). This information was aggregated by year, and economic values were adjusted by the Consumer Price Index \citep{oecd_2017-VV} as:

@@

Data for the qualitative analysis of the social-ecological system were collected at the community--level from official documents used in the creation and designation of the marine reserves \citep{dof_website_2012,dof_website_2013} and based on the authors' experience and knowledge of the communities. These include information on the resource system, the resource units, actors, and the governance system itself (**S1 Table**).

@@

We evaluate the effect that marine reserves have had on four biological and two socioeconomic indicators (Table \ref{table:indicators}). Recall that reserves were implemented to protect lobster and other benthic invertebrates. However, we also use the available fish data, since finfish are targeted in the off-season and because there might be associated cobenefits of the reserve.

```{r}
tibble(Category = c("Biological", "Biological","Biological","Biological", "Socioeconomic", "Socioeconomic"),
           Indicador = c("Lobster density", "Invertebrate density", "Fish biomass", "Fish density", "Income from target species", "Landings from target species"),
       Units = c("org / $m^2$", "org / $m^2$", "Kg / $m^2$", "org / $m^2$", "M MXP", "Metric Tonnes")) %>% 
  knitr::kable(caption = "\\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.", format = "latex", escape = F) %>% 
  kable_styling(latex_options = "HOLD_position")
```

We use a difference-in-differences analysis to evaluate the biological indicators. This approach allows us to estimate the effect that the reserve has on the biological indicators by comparing trends across time and treatments (*i.e.* reserve / control sites \citet{moland_2013-VP,Villasenor-Derbez_2018}). The analysis is performed with a multiple linear regression of the form:

@@

Where year-fixed effects are represented by $\gamma_{it} Year_t$, and $\beta Zone_i$ captures the difference between reserve ($Zone = 1$) and control ($Zone = 0$) sites. The interaction term $\lambda_{it} Year_t\times Zone_i$ represents the mean change in the indicator inside the reserve, for year $t$, with respect to the year of implementation in the control site (See Table \ref{table:com_sum}). When evaluating biomass and densities of the entire benthic or fsh communities, we include $\sigma_j$ to control for species-fixed effects.

@@

\begin{equation}
I = \alpha + \gamma_{t} Year_t + \beta Treated_i + \lambda_{t} Year_t\times Treated_i + \sigma_jCom_j +\epsilon
\label{eqn:soc_reg}
\end{equation}

The model interpretation remains as for Eq \ref{eqn:reg_bio}, but in this case the $Treated$ dummy variable indicates if the community has a reserve ($Treated = 1$) or not ($Treated = 0$) and $\sigma_jCom$ captures community-level fixed-effects. These regressions allows us to make a causal link between the implementation of marine reserves and the observed trends by accounting for temporal and spatial dynamics \citep{depalma_2018}. The effect of the reserve is captured by the $\lambda_t$ coefficient, and represents the difference observed between the control site before the implementation of the reserve and the treated sites at time $t$ after controlling for other time and space variations (*i.e.* $\gamma_t$ and $\beta$ respectively). All model coefficients were estimated via ordinary least-squares and heteroskedastic-robust standard errors \citep{zeileis_2004-7n}. All analyses were performed in R 3.5.0 and R Studio 1.1.453 \citep(R_2018).

@@

The following sections present the effect that marine reserves had on each of the biological and socioeconomic indicators for each coastal community. Results are presented in terms of the difference through time and across sites, relative to the control site on the year of implementation (*i.e.* effect size $\lambda_t$). We also provide an overview of the governance settings of each community, and discuss how these dimensions might be related to the effectiveness of the reserves.

@@

Indicators showed idiosyncratic responses through time for each community. Figure \ref{fig:indicators}A shows positive effect sizes for lobster densities for Isla Natividad and Punta Herrero during the first years, but the effect is eroded through time. These effects are in the order of 0.2 extra organisms $m^{-2}$ but are not significantly different from zero ($p > 0.05$). Lobster densities were only significantly positive for Isla Natividad on the sixth year (*i.e.* 2012; $p < 0.05$), a year after the hypoxia events described by \citet{micheli_2012-EU} caused mass mortality of organisms. Likewise, no changes were detected in fish biomass or invertebrate and fish densities (\ref{fig:indicators}B-D), where effect sizes oscillated around zero without clear trends. Full tables with model coefficients are presented in the supplementary materials (**S2 Table**, **S3 Table**, **S4 Table**).

@@

Lobster landings and revenue were only available for Isla Natividad and Maria Elena (Fig \ref{fig:lobsters}). For all years befor implementation, the effect sizes are close to zero, indicating that the control and treatment sites track each other well and that these are plausible controls. However, effect sizes do not change after the implementation of the reserve. Again, the negative coeficient observed for Isla Natividad on year 5 correspond to the 2011 hypoxia events. The only positive change observed in lobster landings is for Isla Natividad in 2014 ($p < 0.1$). The trhee years of post-implementation data for Maria Elena do not show a significant effect of the reserve. Isla Natividad shows higher revenues after the iumplementation of the reserve, as compared to the control communities. However, these changes are also not ignificantly different and present an increased variation. All regression coefficients for each community and indicator are presented in **S5 Table**.

@@

Although we have little information on the social dimension of these fisheries, we can use the social-ecologycal systems framework (**S1 Table**) to analyze the performance of each governance system (**S6 Table**).

@@

@@

Our results indicate that marine reserves are ineffective in increaseing population sizes of lobsters in the three communities. The socioeconomic indicators pertaining to landings and revenues showed little to no change after reserve implementation. While lobster densities represent an indicator directly tied to reserve objectives, we also evaluate other biological indicators to test for additional effects and fail to detect any effect other then the already reported buffering effect that Natividad reserves can have to environmental variability \citep{micheli_2012-EU}. The lack of expected effectiveness poses the question, why do these communities continue to support the reserves? Understanding the social-ecological context in which these communities and their reserves operate might provide insights to this question. Here, we discuss plausible explanations to lack of effectiveness observed onthe biological and socioeconomic indicators based on our social-ecological system analysis. We also discuss potential shortcomings in our analysis, and provide management recommendations to improve reserve effectiveness.

Our approach to evaluate the temporal and spatial changes of each indicator provides a more robust measure of reserve effectiveness. Some works have solely focused on an inside-outside comparison of indicators \citep{guidetti_2014-8Z,friedlander_2017-oI,rodriguez_2017-PD}, which do not address temporal variability \citep{depalma_2018}. Other works have compared the trend observed within a reserve through time \citep{betti_2017-lq}, which cannot distinguish between the temporal trends in a reserve and the entire system \citep{depalma_2018}. By accounting for trends between sites and through time, we can control for time and space dynamics, and provide a better identification of the effect.

@@

There are two plausible explanations for the observed lack of effectiveness. First, marine reserves are only likely to provide fisheries benefits if initial population sizes are low \citep{hilborn_2006} and are poorly managed. However, both stocks were at some point certified by the Marine Stewardship Council \citep{prezramrez_2016-J1}. The fishery is managed via species-specific minimum catch sizes, seasonal closures, and escapement windows on traps \cite{dof_website_1993}.
A second plausible explanation lies in the design of these areas. Small reserves are unlikely to have a significant effect on population size if they fail to protect a significant portion of it. Intuitvely, small reserves have little or no displacement of fishing effort, and fishers do not perceive the short-term costs associated to the first years of reserve implementation \citep{ovando_2016-Wg}. On either case, these might explain why reserves were not effective, but not why they still receive community support.

While reserves fail to provide conservation or economic benefits, they provide the community with access to funding from the filantrhopic sector. Furthermore, reserves can provide a joint enterprise to bring a community together, which promotes social cohesion and social capital.

@@

Citar Fulton Marine Policy

@@

\paragraph*{S1 Figure}
\label{S1_Figure}
Map of control and treated sites in A and control and treated landings in B

\paragraph*{S2 Figure}
\label{S2_Figure}
Timeseries of indicators for IN

\paragraph*{S3 Figure}
\label{S3_Figure}
Timeseries of indicators for ME

\paragraph*{S4 Figure}
\label{S3_Figure}
Timeseries of indicators for PH
